---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Drone Image Utilities

`uasimg` helps manage images taken from an unmanned aerial vehicle (drone) that have been collected with the intent to stitch into 2D amd 3D products. The package does **not** stitch images, but helps you create catalogs of your image data, examine their locations and properties, export image centroids and estimated footprints as GIS files, and create world files so individual images can be viewed in GIS software.

## Applications

`uasimg` was developed to help with the following common data management tasks:

1. Doing a quick check in the field of photos after a flight, assessing their locations and estimated image overlap. This can help a pilot determine if a flight was successful, or needs to be redone.

1. Subsetting images for further processing with a photogrammetry (stitching) program like Pix4D or Agisoft. Omitting images with an extreme amount of overlap can reduce processing time and improve results.

1. Creating individual Flight Summary pages in HTML, as the backbone of an image catalog.

1. Creating world files for individual drone images, using the image EXIF data to model the ground footprint and rotation, so they can imported in GIS software and appear in their approximate location.

**Note**: image locations and footprints are based on the metadata saved in the image files (e.g., relative altitude above the launch point, camera compass direction). Thus they should taken as *estimates only*.


## Installation

`uasimg` is not on CRAN (yet), but you can install it from GitHub.

*Note*: Windows users need *RTools* to build packages from source files (which is basically what you do when you install from GitHub). RTools is not a R package, rather its a set of utilities that you install separately. You can download and install RTools from <https://cran.r-project.org/bin/windows/Rtools/>. Alternately, you can install RTools from within R by running:

```{r install_rtools, eval=FALSE}
install.packages('installr')
installr::install.Rtools()
```

After RTools is installed, you can install `uasimg` with:

```{r install_devtools, eval=FALSE}
## install.packages("remotes") 
library (remotes)
remotes::install_github("ucanr-igis/uasimg")
```

### Dependent Packages

`uasimg` requires several dependent packages, including *sf*, *dplyr*, *tidyr*, *leaflet*, *htmltools* and others (see the DESCRIPTION file for a complete list). Normally missing packages are installed at the same time, but if you get an error message about a missing package(s), install the dependent packages separately (i.e., from the 'Packages' pane in RStudio) then run `remotes::install_github("ucanr-igis/uasimg", dependencies=FALSE)`. 

### exiftool

To read the EXIF data from the image files, `uasimg` requires an external command line tool called 'exiftool'. exiftool can be installed in four steps:

 1. download the file for your OS from: http://www.sno.phy.queensu.ca/~phil/exiftool/
 1. uncompress / unzip
 1. rename the executable file from *exiftool(-k).exe* to *exiftool.exe*  
     **Note**: if you have file extensions hidden in Windows Explorer, you won't see *.exe* in the filename. In that case, just rename '*exiftool(-k)*' to '*exiftool*'.
 1. move the executable file to a directory on the path (e.g., c:\\windows). Note: putting it in c:\\windows\\system32 does *not* seem to work.

## Supported Cameras

To see a list of known cameras (sensors), run ```uas_cameras()``` with no arguments. If your camera is not listed, you may submit an [issue](https://github.com/ucanr-igis/uasimg/issues) on GitHub to have it added, or pass the camera parameters in as a csv file. For details see the help page (```?uas_cameras```) or contact the package author.

## Data Requirements

Most functions in `uasimg` use location data saved in the EXIF data (header) of image files themselves. This assumes the camera saves the location in the images using a GPS coordinate from the drone or the camera itself. To compute  footprints, the package also needs to know the height at which images were taken. Some drones (including many DJI drones) record the relative flight altitude (above the launch point) in the image file. Flight height can also be entered manually as an argument.

Requirements for `uasimg` include:

 - images must have been taken by a camera that saved the GPS coordinates  
 - image files should be grouped in directories folders (typically each directory containing the images from one flight)
 - the camera model must be one of the ones known by the `uasimg` package (see below)  

Additional requirements to generate estimated footprints:

 - the height above ground level must be saved in the image files, or passed as an argument. If passed as an argument, the assumption is that all images were taken from the same height.  
 - it is presumed that images were taken at nadir (camera pointing straight down)

# Usage Overview

You always start with `uas_info()`, feeding it or more directories of drone images. This function  extracts the image metadata, computes footprints, reads supplemental metadata, etc. 

The object returned by `uas_info()` is not very useful by itself. The results are generally saved to a variable then fed into other functions that do useful things:

- `uas_report()` creates 'Flight Summaries' as standalone HTML pages, with options to create image thumbnails. SAMPLE HERE

- `uas_toc()` creates a Table of Contents page for several Flight Summaries, with options to copy all the catalog files to a single folder so the catalog is in one place.

- `uas_exp_shp()` exports the centroids and footprints of individual images as Shapefiles, as well as the overall flight arrea.

- `uas_worldfile()` creates small external files that allow images to be imported into GIS software and appear in their approximate footprint.


For more info, see the `uasimg` Vignette(s) XXXXXXXXXXXXXXXXX and function help pages.

## Example

The general usage is to first create a metadata object for one or more directories of images using the *uas_info()* function, saving the result to a variable. 

```{r run_uas_info, eval=FALSE}
library(uasimg)
hast_ft1_info <- uas_info("c:/Drone_Projects/Hastings/Flt01_1443_1446_250ft")
```

Once an image collection metadata object has been created, you can generate outputs.

```{r create_rpt, eval=FALSE}
## Generate a flight summary
uas_report(hast_ft1_info)

## Export image centroid, footprints, and minimum convex polygon as Shapefiles
uas_exp_shp(hast_ft1_info)

## Generate estimated world files so the images can be imported into ArcGIS or QGIS
uas_worldfile(hast_ft1_info)
```

## World Files

(**MOVE TO A VIGNETTE**)

Drone images typically save the coordinates of the camera, but do not include the width, length, or compass angle. A "world file" is a small external file that contains these additional parameters, so that when you import an individual image into a GIS program like ArcGIS or QGIS, the image will appear in its approximate footprint on the ground. 

You can create world files, readable by ArcGIS and QGIS, with `uas_worldfile()`. `uas_worldfile()` can create three types of world files, including `aux.xml`, `jpw` and `tfw`, and `prj` files. `aux.xml` is the most recognized format and the default. See the function help page for details.

# Bugs, Questions, and Feature Requests

To report a bug, add your camera to the package, or suggest a new feature, please create an [issue](https://github.com/ucanr-igis/uasimg/issues) on GitHub, or contact the package author.
